Solar wind, F10.7, and geomagnetic activity relationship to the equatorial plasma mass density at geosynchronous orbit

### 2016-08-17 paper/paper.tex
Misc edits

\begin{document}
\title{Solar wind and geomagnetic activity influence on the equatorial plasma mass density at geosynchronous orbit}
\author{V. Veibell and R.S. Weigel}
\doublespacing
\maketitle

@@

%\twocolumn[
%  \begin{@twocolumnfalse}
%\maketitle
%\hrule
\begin{abstract}
We consider two types of events, identified by decreases in $D_{st}$ below a threshold value and increases in the the equatorial mass density at geosynchronous altitudes, $\rho_{eq}$, above a threshold value using the \cite{Takahashi2010} dataset.  From the $D_{st}$ events and 1-day averages, we find that there is a statistically weak and small-amplitude difference between $\rho_{eq}$ on the day of the event and the days before and after.

For hourly averages, $\rho_{eq}$ following the onset of a $D_{st}$ event does not depend on the north-south component of the interplanetary magnetic field, $B_z$ near onset time. The primary factor that determines the post-$D_{st}$ event onset peak amplitude in $\rho_{eq}$ in the hourly averages is elevated $F10.7$.  From the $\rho_{eq}$ events, we find a weak dependence on $B_{z}$ after the onset of an event, with higher average $B_{z}$ four hours after the event onset corresponding to larger $\rho_{eq}$ values 24-36 hours after onset.  In contrast, the average of $B_z$ four hours prior to $\rho_{eq}$ events does not have a statistically significant impact on $\rho_{eq}$ after onset.
\end{abstract}
%\vspace{2em}
%\hrule
%\vspace{2em}
% \end{@twocolumnfalse}
%  ]

@@

This work presents an analysis of two types of events intended to provide insight into the relationship between $\rho_{eq}$ and solar wind and geomagnetic parameters by considering both large geomagnetic events and large $\rho_{eq}$ events.  The analysis attempts to address a major difficulty with the identification of the cause in $\rho_{eq}$ enhancements, which can occur for two reasons: (1) an extended period of positive $B_z$, which leads to low geomagnetic activity, plasmasphere refilling, and a possible expansion of the plasmasphere to geosynchronous altitudes and (2) strong negative $B_z$, which (a) leads to the inward motion of the higher density plasmasphere and leaves geosynchronous altitudes in the lower density plasma trough just outside of the plasmapause and (b) causes possible magnetospheric processes to enhance the mass density at geosynchronous altitudes.

On average, around the time of large geomagnetic events, there is a statistically weak enhancement of daily-averaged $\rho_{eq}$. For hourly averages $\rho_{eq}$ shows enhancements over a 6-hour window that contains too few events for a statistical analysis.  When the constraint of extended geomagnetic events (greater than 12 hours below a threshold value) and the MLT of the observation of $\rho{eq}$ (06:00-12:00) is removed to allow for more events, statistically significant variations are found, with a peak occurring approximately six hours after the event onset.  For the hourly-averaged measurements, the average of $B_{z}$ around onset time did not have a significant impact on $\rho_{eq}$ after onset.

@@

For $\rho_{eq}$, more positive $B_{z}$ after onset lead to higher values of subsequent $\rho_{eq}$, which indicates that a substantial portion of the events may due to the apparent mass refilling mechanism described by \cite{Denton2016}.  Without additional information about the location of the plasmapause, it may not be possible to separate out the conditions required for enhancement due to enhanced magnetospheric activity from this mechanism.

>>>>

\begin{document}
\title{Solar wind, F10.7, and geomagnetic activity relationship to the equatorial plasma mass density at geosynchronous orbit}
\author{V. Veibell and R.S. Weigel}
\doublespacing
\maketitle

@@

%\twocolumn[
%  \begin{@twocolumnfalse}
%\maketitle
%\hrule

% Hourly averages are required to show a statistically significant mass density response during high geomagnetic activity
% The response of mass density for high geomagnetic activity intervals is strongest when F10.7 is elevated
% There is no solar wind variation signature associated with large mass density events

\begin{abstract}
We consider two types of events, identified by decreases in $D_{st}$ below a threshold value and increases in the the equatorial mass density at geosynchronous altitudes, $\rho_{eq}$, above a threshold value using the \cite{Takahashi2010} dataset.  From the $D_{st}$ events and 1-day averages, we find that there is a statistically weak and small-amplitude difference between $\rho_{eq}$ on the day of the event and the days before and after.  When hourly averages are considered, a signficant peak is found to occur six hours after event onset and the primary factor that determines the post-onset peak amplitude in $\rho_{eq}$ in the hourly averages is elevated $F10.7$.  In addition, for hourly averages, $\rho_{eq}$ following the onset of a $D_{st}$ event does not depend on the north-south component of the interplanetary magnetic field, $B_z$ near the time of onset.  From the $\rho_{eq}$ events, we find a weak dependence on $B_{z}$ after the onset of an event, with higher average $B_{z}$ four hours after the event onset corresponding to larger $\rho_{eq}$ values 24-36 hours after onset.  In contrast, the average of $B_z$ four hours prior to $\rho_{eq}$ events does not have a statistically significant impact on $\rho_{eq}$ after onset.
\end{abstract}
%\vspace{2em}
%\hrule
%\vspace{2em}
% \end{@twocolumnfalse}
%  ]

@@

This work presents an analysis of two types of events intended to provide insight into the relationship between $\rho_{eq}$ and solar wind and geomagnetic parameters by considering both large geomagnetic events and large $\rho_{eq}$ events.  The two types of events were considered in order to account for a difficulty with the identification of the cause in $\rho_{eq}$ enhancements, which can occur for two reasons: (1) an extended period of positive $B_z$, which leads to low geomagnetic activity, plasmasphere refilling, and a possible expansion of the plasmasphere to geosynchronous altitudes and (2) strong negative $B_z$, which (a) leads to the inward motion of the higher density plasmasphere and leaves geosynchronous altitudes in the lower density plasma trough just outside of the plasmapause and (b) causes possible magnetospheric processes to enhance the mass density at geosynchronous altitudes.

On average, around the time of large geomagnetic events, there is a statistically weak enhancement of daily-averaged $\rho_{eq}$ measurements. For hourly averages, $\rho_{eq}$ shows enhancements over a 6-hour post-onset window that contains too few events for a statistical analysis.  When the constraint of extended geomagnetic events (greater than 12 hours below a threshold value) and the MLT of the observation of $\rho_{eq}$ (06:00-12:00) is removed to allow for more events, statistically significant variations are found, with a peak occurring approximately six hours after the event onset.  For the hourly-averaged measurements, the average of $B_{z}$ around onset time did not have a significant impact on $\rho_{eq}$ after onset.

@@

For $\rho_{eq}$, more positive $B_{z}$ after onset corresponded to higher values of subsequent $\rho_{eq}$, which indicates that a substantial portion of the events may due to the apparent mass refilling mechanism described by \cite{Denton2016}.  In order to separate this mechanism from an enhancement due to magnetospheric activity, additional information about the location of the plasmapause may be required.
